---
title: "Final project"
output:
  word_document:
    fig_height: 5
    fig_width: 8
date: "12/6/2024"
author: "Jared Adam"
---

# Packages 

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE)
options(show.signif.stars = FALSE)

library(lme4)
library(ggplot2)
library(ggthemes)
library(tidyverse)
library(car)
library(effects)
library(mosaic)
library(ggmosaic)
library(ggResidpanel)
library(catstats2)
library(effects)
library(gtsummary)
library(modelsummary)
library(ggpp)
library(multcomp)
library(emmeans)
library(MuMIn)
library(DiagrammeR)
library(DiagrammeRsvg)
# theme_set(theme_bw()) #Prevents need for + theme_bw() in ggplots
```

# Data import and wrangling

```{r}
set.seed(654321)
s21 <- read_csv('data/2021 Sentinel Prey Assessment.csv')
s22 <- read_csv("data/PSA_CE2_SentinelPrey.csv")
s23 <- read_csv('data/PSA_Sent.prey.2023.csv')

# I need to get total predation into a column as a binary. 1 = predation, 0 = not

# 2021 cleaning 
s21
clean21 <- s21 %>% 
  mutate(year = '2021') %>% 
  dplyr::select(location, year, growth_stage, plot_id, rep.block, treatment, to.predated) %>% 
  mutate(to.predated = as.double(to.predated)) %>% 
  dplyr::rename(block = rep.block) %>% 
  group_by(location, year, growth_stage, plot_id, block, treatment) %>% 
  # dplyr::summarise(total = sum(to.predated)) %>% 
  na.omit() %>% 
  mutate(treatment = case_when(
    treatment == '33' ~ '3',
    .default = as.factor(treatment))) %>% 
  dplyr::filter(treatment != '6',
                treatment != '7',
                treatment != '8') %>% 
  mutate_at(vars(1:6), as.factor) 


# 2022 cleaning 
s22
unique(s22$treatment)
unique(s22$growth_stage)
clean22 <- s22 %>% 
  mutate(year = '2022') %>% 
  dplyr::select(location, year, growth_stage, plotid, block, treatment, to.predated) %>% 
  dplyr::rename(plot_id = plotid) %>% 
  mutate(growth_stage = case_when(growth_stage == 'R2' ~ 'R3',
                                  .default = as.character(growth_stage))) %>% 
  dplyr::group_by(location, year, growth_stage, plot_id, block, treatment) %>% 
  # dplyr::summarise(total = sum(to.predated)) %>% 
  mutate_at(vars(1:6), as.factor)


# 2023 cleaning 

clean23 <- s23 %>%
  mutate(year = '2023') %>%
  relocate(am.partial, am.absent, pm.partial, pm.absent) %>% 
  mutate_at(vars(1:4), as.double) %>% 
  mutate(to.predated = if_else(am.partial | am.absent | pm.partial | pm.absent == 1, 1, 0)) %>% 
  relocate(to.predated)%>%  
  mutate(growth_stage = case_when((location == 'NC' & date == '7/20/2023') ~ 'R3',
                                  .default = as.character(growth_stage))) %>%  
  dplyr::select(location, year, growth_stage, plotid, block, treatmetn, to.predated) %>% 
  dplyr::rename(plot_id = plotid, 
         treatment = treatmetn) %>%
  group_by(location, year, growth_stage, plot_id, block, treatment) %>% 
  na.omit() %>% 
  filter(treatment != 5) %>% 
  mutate_at(vars(1:6),as.factor)

# and in the darkness, bind them 
sent <- rbind(clean21, clean22, clean23)
as_tibble(sent)
dim(sent)
unique(sent$location)

```


# Experimental design 

Data: Sentinel Prey assessment of arthropod-predator activity in corn fields. 

These data come from the Precision Sustainable Agriculture effort through the USDA. I am the lead on the entomology component of this project and responsible for analyzing this three year data set which spans multiple states. This effort began during my Master's degree, but I only analyzed Pennsylvania data for my thesis. 

**Study design: **
**Treatments** = 4; No cover crop, early-terminated cover crop, late-terminated cover crop, planting green
**Plots** = 20; 5-6 blocks (location dependent) composed of 4 plots each = 20 plots / study site / year
**Years** = 3 (2021,2022,2023)
**Locations** = This project comprises 16 states. Not all states collected sentinel prey data every year. Each site year was in a different field. 
**Effort** = Data were collected at three corn growth stages / year (V3,V5,R3).
**Sample** = 6 sentinel prey traps were placed in each plot = 120 samples collected / growth stage. Total sample effort per state per season = 360 samples. 

**Variables:**
**Response** = Total level of predation. This is a binomial of 6 traps/ counts per plot. Pseudoreplication is account for in the random term. 
**Explanatory** = Crop growth stage (timing, three levels) and treatment (four levels). I am not interested in the fixed effects of location. 
**Random effects** = Plot in block in location, in year. I want to account for pseduoreplication and all of the site/year combinations.

Plots were randomly assigned to each block. Field sites were as random as they could be at each respective research station. Sentinel prey traps were placed between pre-determined rows and at specific length intervals within each plot to maintain consistency. 

# Locating missing values

```{r}
library(mi)
# make an object of the missing df and then present the image
tdf <- missing_data.frame(data.frame(sent))
image(tdf)
```

# Preliminary data display 

```{r}
sent %>% 
  ggplot() +
  geom_mosaic(aes(x = product(treatment), fill = to.predated), offset = 0.04)+
  facet_wrap(~growth_stage)+
  scale_fill_colorblind()+
  labs(title = 'Mosaic plot of Predation by Treatment and Growth Stage',
       y = 'Density',
       x = ' Treatment')


```


# Modeling 

Theoretical model:

${Predation|Treatment,GrowthStage} \sim \beta_0 + \beta_1I_{Treatment = 2}_i +  \beta_2I_{Treatment = 4}_i + \beta_3I_{Treatment = 3}_i + \beta_4I_{GrowthStage = V5}_i + \beta_5I_{GrowthStage = V3}_i + \Year_i + \Location_i + \Block_i + \Plot_i + \Error_i$

Add the appropriate stuff below

$\pi\{Predation|Treatment, GrowthStage} \sim \beta_0 + \beta_1I_{Treatment = 2} + \beta_2I_{Treatment = 4} + \beta_3I_{Treatment = 3} + \beta_4I_{GrowthStage = V5} + \beta_5I_{GrowthStage = V3}$


## Initial model 

```{r}
sent

m2 <- glmer(to.predated ~ treatment*growth_stage + (1|year/location/block/plot_id) , family = binomial, data = sent)
summary(m2)

Anova(m2)

plot(allEffects(m2), type = 'link',ylab = 'estimated log differnces in total predation', grid = T)
plot(allEffects(m2), type = 'response', xlim = c(0,1),ylab = 'estimated probability differnces in total predation', grid = T)



```
* There is weak evidence against the null of no interaction between treatment and growth stage Chi squared(6) = 8.25, p = 0.22, after controlling for the random effect of plot nested in block nested in location nested in year, and will remove the interaction term from the model. 

## Model refinement

```{r}
m3 <- glmer(to.predated ~ treatment+growth_stage + (1|year/location/block/plot_id) , family = binomial, data = sent)
summary(m3)

Anova(m3)

plot(allEffects(m3), type = 'link', ylab = 'Estimated log differnces in total predation', grid = T)
plot(allEffects(m3), type = 'response', xlim = c(0,1),ylab = 'Estimated probability differnces in total predation', grid = T)


cld(emmeans(m3, ~treatment, adjust = 'tukey', type = 'response'), Letters = letters)
cld(emmeans(m3, ~growth_stage, adjust = 'tukey', type = 'response'), Letters = letters)

```
### R squared and ICC

```{r}
r.squaredGLMM(m3)
library(performance)
r2(m3)

summary(m3)

# for plot 
(2.600e-01) / (2.600e-01 + 1.700e-01 + .672e+00 + 5.569e-06)


```
The ICC for plot id is 0.236.

# Final plots 

```{r}
sent %>% 
  ggplot()+
  geom_mosaic(aes(x = product(treatment), fill = to.predated))+
  scale_fill_manual(values = c("#E7298A","#1B9E77"))+
  scale_x_productlist(labels=c("No CC", "Early", "Late", "Green"))+
  labs(title = 'Whole Team Sentinel Prey ~ Treatment',
       x = 'Treatment',
       y = 'Estimated proportion',
       caption = 'Figure 5:\nTotal predation binary where 1 = predation and 0 = no predation.\nPredation levels differed between the no-cover and late-terminated treatment\nand the no-cover and planting-green treatment (p < 0.05).\nCompact letter display denotes differences among treatments.')+
  guides(fill = guide_legend(title = 'Predation:\n0 = no predation\n1 = predation'))+
  theme(legend.title = element_text(size = 14),
        legend.text = element_text(size = 14),
        axis.text.x = element_text(size=16),
        axis.text.y = element_text(size = 16),
        axis.title = element_text(size = 18),
        plot.title = element_text(size = 18),
        plot.subtitle = element_text(size = 12), 
        panel.grid.major.y = element_line(color = "darkgrey"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        strip.text = element_text(size = 12),
        plot.caption = element_text(hjust = 0, size = 12, color = "grey25"),
        axis.ticks.length=unit(.25, "cm"))+
  annotate('text', x = 0.15, y = 0.95, label = 'a', size = 8)+
  annotate('text', x = .4, y = 0.95, label = 'a', size = 8)+
  annotate('text', x = .65, y = 0.95, label = 'b', size = 8)+
  annotate('text', x = .875, y = 0.95, label = 'c', size = 8)


sent %>% 
  ggplot()+
  geom_mosaic(aes(x = product(growth_stage), fill = to.predated))+
  scale_fill_manual(values = c("#E7298A","#1B9E77"))+
  labs(title = 'Whole Team Sentinel Prey ~ Growth Stage',
       x = 'Growth Stage',
       y = 'Esimiated proportion',
       caption = 'Figure 6:\nTotal predation binary where 1 = predation and 0 = no predation.\nPredation levels differed among all growth stages (p < 0.05), where R3 was the highest,\nfollowed by V5, and then V3. Growth stages are ordered in reverse-chronological order.\nCompact letter display denotes differences among growth stages')+
  guides(fill = guide_legend(title = 'Predation:\n0 = no predation\n1 = predation'))+
  theme(legend.title = element_text(size = 14),
        legend.text = element_text(size = 14),
        axis.text.x = element_text(size=16),
        axis.text.y = element_text(size = 16),
        axis.title = element_text(size = 18),
        plot.title = element_text(size = 18),
        plot.subtitle = element_text(size = 12), 
        panel.grid.major.y = element_line(color = "darkgrey"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        strip.text = element_text(size = 12),
        plot.caption = element_text(hjust = 0, size = 12, color = "grey25"),
        axis.ticks.length=unit(.25, "cm"))+
  annotate('text', x = 0.175, y = 0.95, label = 'a', size = 8)+
  annotate('text', x = .5, y = 0.95, label = 'b', size = 8)+
  annotate('text', x = .85, y = 0.95, label = 'c', size = 8)

```

# Table 1

```{r}
# table as a proportion 
sent %>% 
  group_by(location, treatment, growth_stage) %>% 
  summary()

sent %>% 
  group_by(growth_stage) %>% 
  summary()


tally(to.predated ~ treatment + growth_stage, data = sent)

datasummary_balance( ~ to.predated,data = sent)

# install.packages('openxlsx')
# library(openxlsx)
# 
# table <- favstats(to.predated ~ treatment + growth_stage, data = sent)
# write.xlsx(table, 'test.xlsx')
# favstats(to.predated ~ growth_stage, data = sent)
# favstats(to.predated ~ treatment, data = sent)
# citation('mosaic')

```



# Study design figure 

```{r}
tally(~ year, sent)
tally(~location, sent)
tally(~block, sent)
tally(~plot_id, sent)

```

```{r}
names1 <- list("Year 2021", "Year 2022", "Year 2023")
names2 <- list("Loc. 1", "...", "Loc. 16", "...", "Loc. 1", "...", "Loc. 16")
names3 <- list("Block 1", "...", "Block 6")
names4 <- list("Plot 1", "...", "Plot 4")
namesF <- list("Treatment", "Growth Stage")
namesR <- list("Year", "Location in Year", "Block in Location", "Observation Error")
```


```{r echo = F, fig.height = 12, fig.width = 10}
textforgrViz <- "digraph boxes_and_circles {graph [rankdir = LR, overlap = true, fontsize = 10]; #rankdir = TB also of interest
  node [shape = box,
        fontname = Helvetica];  A [label = '@@3'];   B [label = '@@2'];  C [label = '@@1']; 
  node [shape = box];
    FixedLevel1 [label = '@@57', alpha = 0.2,
        fontcolor = blue];   
    FixedLevel2 [label ='@@58', alpha = 0.2, 
        fontcolor = blue];
  node [shape = box];
    RandomLevel1 [label = '@@59', alpha = 0.2,
        fontcolor = darkgreen];
    RandomLevel2 [label = '@@60', alpha = 0.2,
        fontcolor = darkgreen];
    RandomLevel3 [label = '@@61', alpha = 0.2,
        fontcolor = darkgreen];
    RandomLevel4 [label = '@@62', alpha = 0.2,
        fontcolor = darkgreen];
  node [shape = circle,
        fixedsize = true,
        width = 0.9] ;  B1 [label = '@@4']; B2 [label = '@@5']; B3 [label = '@@6']; B4 [label = '@@7']; B5 [label = '@@8']; B6 [label = '@@9']; B7 [label = '@@10'];
    node [shape = circle,
        fixedsize = true,
        width = 0.7];  C1 [label = '@@11']; C2 [label = '@@12']; C3 [label = '@@13']; C4 [label = '@@14']; C5 [label = '@@15']; C6 [label = '@@16']; C7 [label = '@@17']; C8 [label = '@@18']; C9 [label = '@@19']; C10 [label = '@@20']; C11 [label = '@@21']; C12 [label = '@@22']; C13 [label = '@@23']; C14 [label = '@@24']; C15 [label = '@@25'];
    node [shape = circle, 
          fixedsize = true,
          width = 0.7]; D1 [label = '@@26']; D2 [label = '@@27']; D3 [label = '@@28']; D4 [label = '@@29']; D5 [label = '@@30']; D6 [label = '@@31']; D7 [label = '@@32']; 
          D8 [label = '@@33']; 
          D9 [label = '@@34']; D10 [label = '@@35']; D11 [label = '@@36']; D12 [label = '@@37']; D13 [label = '@@38']; D14 [label = '@@39']; D15 [label = '@@40'];
          D16 [label = '@@41'];
          D17 [label = '@@42']; D18 [label = '@@43']; D19 [label = '@@44']; D20 [label = '@@45']; D21 [label = '@@46']; D22 [label = '@@47']; D23 [label = '@@48'];
          D24 [label = '@@49'];
          D25 [label = '@@50']; D26 [label = '@@51']; D27 [label = '@@52']; D28 [label = '@@53']; D29 [label = '@@54']; D30 [label = '@@55']; D31 [label = '@@56'];
          
  # several 'edge' statements
  A -> {B1 B2 B3}; B -> B4; C -> {B5 B6 B7}; 
  B1 -> {C1 C2 C3}; B2 -> C4; B3 -> {C5 C6 C7}; B4 -> C8; B5 -> {C9 C10 C11}; B6 -> C12; B7 -> {C13 C14 C15};
  C1 -> {D1 D2 D3}; C2 -> D4; C3 -> {D5 D6 D7}; C4 -> D8; C5 -> {D9 D10 D11}; C6 -> D12; C7 -> {D13 D14 D15}; C8 -> D16; C9 -> {D17 D18 D19}; C10 -> D20; C11 -> {D21 D22 D23}; C12 -> D24; 
  C13 -> {D25 D26 D27}; C14 -> D28; C15 -> {D29 D30 D31};
  FixedLevel1 -> FixedLevel2 [color = white];
  RandomLevel1 -> RandomLevel2 -> RandomLevel3 -> RandomLevel4 [color = white]} \n
  [1]: names1[1];
  [2]: names1[2];
  [3]: names1[3];
  [4]: names2[1];
  [5]: names2[2];
  [6]: names2[3];
  [7]: names2[4];
  [8]: names2[5];
  [9]: names2[6];
  [10]: names2[7];
  [11]: names3[1];
  [12]: names3[2];
  [13]: names3[3];
  [14]: names3[2];
  [15]: names3[1];
  [16]: names3[2];
  [17]: names3[3];
  [18]: names3[2];
  [19]: names3[1];
  [20]: names3[2];
  [21]: names3[3];
  [22]: names3[2];
  [23]: names3[1];
  [24]: names3[2];
  [25]: names3[3];
  [26]: names4[1];
  [27]: names4[2];
  [28]: names4[3];
  [29]: names4[2];
  [30]: names4[1];
  [31]: names4[2];
  [32]: names4[3];
  [33]: names4[2];
  [34]: names4[1];
  [35]: names4[2]; 
  [36]: names4[3];
  [37]: names4[2];
  [38]: names4[1];
  [39]: names4[2];
  [40]: names4[3];
  [41]: names4[2];
  [42]: names4[1];
  [43]: names4[2];
  [44]: names4[3];
  [45]: names4[2]; 
  [46]: names4[1];
  [47]: names4[2];
  [48]: names4[3];
  [49]: names4[2];
  [50]: names4[1];
  [51]: names4[2];
  [52]: names4[3];
  [53]: names4[2];
  [54]: names4[1];
  [55]: names4[2];
  [56]: names4[3];
  [57]: namesF[1];
  [58]: namesF[2];
  [59]: namesR[1];
  [60]: namesR[2];
  [61]: namesR[3];
  [62]: namesR[4];
"

g3 <- textforgrViz %>% grViz()

g3 %>%
  DiagrammeRsvg::export_svg() %>%
  charToRaw() %>%
  rsvg::rsvg_png("threeleveld.png")
```


# Citations

```{r}
citation('DiagrammeRsvg')
citation('car')
citation('effects')
citation('emmeans')
citation('performance')
```


