---
title: "512 Project Part I"
output:
  word_document:
    fig_height: 5
    fig_width: 8
date: "Due Sept 27"
author: "Jared Adam"
---
**The number of conflicts with dplyr from all of the packages we must download is becoming annoying and breaking code.** 

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE)
options(show.signif.stars = FALSE)

library(lme4)
library(ggplot2)
library(ggthemes)
library(tidyverse)
library(car)
library(effects)
library(mosaic)
# library(ggResidpanel)
library(catstats2)
library(effects)
library(gtsummary)
theme_set(theme_bw()) #Prevents need for + theme_bw() in ggplots
```

# Part I (512 only, project proposal, 25 pts): 

1) Read in your data set and run `dim` on it:

```{r}
set.seed(654321)
s21 <- X2021_Sentinel_Prey_Assessment
s22 <- PSA_CE2_SentinelPrey
s23 <- PSA_Sent_prey_2023

# I need to get total predation into a column as a binary. 1 = predation, 0 = not

# 2021 cleaning 
s21
clean21 <- s21 %>% 
  mutate(year = '2021') %>% 
  dplyr::select(location, year, growth_stage, plot_id, rep.block, treatment, to.predated) %>% 
  mutate(to.predated = as.double(to.predated)) %>% 
  dplyr::rename(block = rep.block) %>% 
  group_by(location, year, growth_stage, plot_id, block, treatment) %>% 
  # dplyr::summarise(total = sum(to.predated)) %>% 
  na.omit() %>% 
  mutate(treatment = case_when(
    treatment == '33' ~ '3',
    .default = as.factor(treatment))) %>% 
  dplyr::filter(treatment != '6',
                treatment != '7',
                treatment != '8') %>% 
  mutate_at(vars(1:6), as.factor) 
# %>% 
#   mutate(total = case_when(total == 7 ~ 6,
#                            .default = as.numeric(total)))
unique(clean21$treatment)
which(clean21$total>6)
# 2022 cleaning 
s22
unique(s22$treatment)
unique(s22$growth_stage)
clean22 <- s22 %>% 
  mutate(year = '2022') %>% 
  dplyr::select(location, year, growth_stage, plotid, block, treatment, to.predated) %>% 
  dplyr::rename(plot_id = plotid) %>% 
  mutate(growth_stage = case_when(growth_stage == 'R2' ~ 'R3',
                                  .default = as.character(growth_stage))) %>% 
  dplyr::group_by(location, year, growth_stage, plot_id, block, treatment) %>% 
  # dplyr::summarise(total = sum(to.predated)) %>% 
  mutate_at(vars(1:6), as.factor)

# 2023 cleaning 

# all of the below notes can be ignored for now. This was with total predation for the summary statistics, not the binary cleaning for the model
# some of my totals are adding up to be greater than 6... Why is this? States in question: OH and NC?
# I found these errors after looking at the enhanced strip chart
# let's pull these states out and have a closer look
# 9/17/2024: I cannot do anything with these until I regain access to the Google drive. These are fucked. NC appears to have made their own sampling regime up. 
# possible solution: If NC and OH have total values >6, I can change them based on the ratio. E.g., pull states out, get total number of samples record total/n... 
erroneous_df <- s23 %>% 
  dplyr::filter(location %in% c('OH', 'NC')) %>% 
  arrange(location, growth_stage, plotid) %>% 
  distinct()


colnames(s23)
unique(s23$growth_stage)

clean23 <- s23 %>%
  # dplyr::filter(location != c('OH', 'NC')) %>% 
  mutate(year = '2023') %>%
  relocate(am.partial, am.absent, pm.partial, pm.absent) %>% 
  mutate_at(vars(1:4), as.double) %>% 
  mutate(to.predated = if_else(am.partial | am.absent | pm.partial | pm.absent == 1, 1, 0)) %>% 
  relocate(to.predated)%>% 
  dplyr::select(location, year, growth_stage, plotid, block, treatmetn, to.predated) %>% 
  dplyr::rename(plot_id = plotid, 
         treatment = treatmetn) %>% 
  group_by(location, year, growth_stage, plot_id, block, treatment) %>% 
  # dplyr::summarise(total = sum(to.predated)) %>% 
  na.omit() %>% 
  filter(treatment != 5) %>% 
  mutate_at(vars(1:6),as.factor)
unique(clean23$treatment)
# which(clean23$total >6)
# these all look good, now let's add the nc and oh back in
# rbind(clean23, erroneous_df)

# and in the darkness, bind them 
sent <- rbind(clean21, clean22, clean23)
as_tibble(sent)
dim(sent)

```


2) Prepare a short description of your data set (source if published paper exists), especially providing the study design, sample size, and variables of primary interest. If there is random sampling, note the population sampled from. If there is random assignment, note how and for which variable(s).

Data: Sentinel Prey assessment of arthropod-predator activity in corn fields. 

These data come from the Precision Sustainable Agriculture effort through the USDA. I am the lead on the entomology component of this project and responsible for analyzing this three year data set which spans multiple states. This effort began during my Master's degree, but I only analyzed Pennsylvania data for my thesis. 

**Study design: **
**Treatments** = 4; No cover crop, early-terminated cover crop, late-terminated cover crop, planting green
**Plots** = 20; 5 blocks composed of 4 plots each = 20 plots / study site / year
**Years** = 3 (2021,2022,2023)
**Locations** = This project comprises 16 states. Not all states collected sentinel prey data every year. 
**Effort** = Data were collected at three corn growth stages / year (V3,V5,R3).
**Sample** = 6 sentinel prey traps were placed in each plot = 120 samples collected / growth stage. Total sample effort per state per season = 360 samples. 

**Variables:**
**Response** = Total level of predation. This is transformed into a proportion over a constant total. Binomial response
**Explanatory** = Crop growth stage (timing, three levels) and treatment (four levels). 
**Random effects** = Block and location? I am unsure how to use location. I am not interested in seeing how each state differs in the analysis because that is *not* a research question. I plan to go through and conduct each state's own analysis later.
**Repeated measure** = I visit the same trap three times throughout the year. This feels like a term I should identify. If I were to choose, I would select *growth stage*. 

Plots were randomly assigned to each block. Field sites were as random as they could be at each respective research station. Sentinel prey traps were placed between pre-determined rows and at specific length intervals within each plot to maintain consistency. 

3) Make a `missing_data.frame` plot of your data set and explain any missing values indicated:

```{r}
library(mi)
# make an object of the missing df and then present the image
tdf <- missing_data.frame(data.frame(sent))
image(tdf)

```



4) Discuss any other use in classes or theses for the data - either that you have used it for or are currently working on for future submissions.

**I am working on this for a publication. The Pennsylvania-exclusive data were analyzed for me master's thesis. The larger, whole-project data set is for publication.**

5) Provide at least one display of the data, focusing on the response of interest versus a predictor. If you have multiple predictors, try to plot the response versus those too.

```{r}
sent
# I do not this plotting function for binary data
enhanced_stripchart(to.predated ~ year, data = sent)
enhanced_stripchart(to.predated ~ treatment, data = sent) # 3 = planting green
enhanced_stripchart(to.predated ~ growth_stage, data = sent) # expect higher predation later in the season
enhanced_stripchart(to.predated ~ location, data = sent)
```

**Important to note: all values have a max of 6. These figures as of 9/17/2024 exclude NC and OH from 2023.**

6) Provide an initial model you hope to fit (does not need to be fit). If you fit a model, add a model summary and effects plot.

```{r}
sent

fullm1 <- glmer(to.predated ~ treatment*growth_stage + (growth_stage|year/location/block/plot_id), family = binomial, data = sent) 
summary(fullm1)
hist(residuals(fullm1))
plot(allEffects(fullm1))

```

7) Start to work on a Table 1 that summarizes variables of interest, possibly by groups of interest. At a minimum, summarize the response variable, by a grouping variable if one exists. 

```{r}
# table as a proportion 
all_locs <- sent %>% 
  group_by(location, treatment, growth_stage) %>% 
  dplyr::summarise(prop = mean(to.predated),
                   sd = sd(to.predated),
                   n = n(), 
                   se = sd/sqrt(n))

by_trt <- sent %>% 
  group_by(treatment) %>% 
  dplyr::summarise(prop = mean(to.predated),
                   sd = sd(to.predated),
                   n = n(), 
                   se = sd/sqrt(n))

by_growth <- sent %>% 
  group_by(growth_stage) %>% 
  dplyr::summarise(prop = mean(to.predated),
                   sd = sd(to.predated),
                   n = n(), 
                   se = sd/sqrt(n))


```


8) Provide the names of feedback group members and the date, time, and location of your feedback session interaction.



Graded for completion/not but there are points for participation in a feedback session. Note that 412 students get full credit for this.
